\noindent

Week 7

Attendance:
All group members attended the meeting.

Discussion:

We began by going through the assignment brief together to understand both individual and group parts. The group agreed to split the visualisation work into three main data types: high-dimensional data, graph data, and dynamic data. Two members were assigned to each part. We then explored the full crash dataset and listed the main columns available. Everyone discussed which features seemed most usable and relevant. Some ideas included comparing gender and time of crash, building networks using location and crash counts, and tracking how crashes changed across months. We also checked for missing values and inconsistencies, and noted which columns would need cleaning. Most of the time was spent looking at the structure of the dataset and seeing how it could be divided up in a way that matched the assignment goals.


Plan:

Each group of two will focus on one data type and continue exploring their subset. By next week, they will decide on final variables and start testing small samples. The goal is to have a clear plan for how to visualise each type and what the main question will be. Everyone will also clean their data section and write notes about any issues they find.

\noindent

Week 8

Attendance:
All six members joined the meeting as scheduled.

Discussion:

Each pair reported back on their chosen data type. The group working on high-dimensional data confirmed they would use gender, age, crash time, and speed limit. The graph data group selected region and month, aiming to build a network of crash frequency over time. The dynamic group decided to use weekday and remoteness level to show how patterns shift month by month. We looked at how to avoid overlapping subsets and made sure each part used unique dimensions. We also agreed that each group would be responsible for its full section, from processing to visualisation to writing. A shared folder was created to store all files, including datasets, code, screenshots, and animations. We briefly discussed visual styles and tools, such as using parallel coordinates for high-dimensional data, layout algorithms for graphs, and animated formats for dynamic networks.

Plan:

All groups will begin their first draft visualisation in the coming week. Each team will also take notes while building visuals so they can explain their design choices later. By Week 9, each group should have at least one working version of their graph or plot and be ready to explain what insights they found.